Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2022-06-18 analysis/paper_draft2.Rmd
changed paths for figures; edits to text

load_path <- here::here("analysis","figures","pp15_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("analysis","figures","pp10_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("analysis","figures","pp6_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("analysis","figures","C95plot.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("analysis","figures","simultaneous_sims_plot.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot1.png")

@@

load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot2.png")

@@

load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot3.png")

@@

load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot4.png")

@@

In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ that is selected. For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), we should let $\alpha = 1$ (equal weight between current and historical data). This will maximize the model's ability to detect bias, even when the number of strikes is small. This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes. Higher bias detection rates are obtained for higher values of alpha, acorss compatibility scenarios (Fig X). This effect is most pronounced for scenarios with a low number of strikes and few historical trails (top left paned of fig X). There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X). Overall, though, increasing $\alpha$ will increase bias detection under current and histrotical trial compatiblity.

@@

Finally, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials. This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof. Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias. @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior. The sensitivity to the initial prior’s SD, however, is reduced by the inclusion of historical, especially when $\alpha$ is close to 1. Fig X shows the effect of the initial prior’s SD on bias detection rate under a number of compatibility scenarios and alpha values. We can see that the effect of setting the initial prior SD to 1 (half the value used in our simulations) is most pronounced when $\alpha$ is 0.1, i.e., borrowing less information from the historical data. As alpha increases, the impact of lowering the initial prior SD is reduced. Fig X also illustrates that increasing the prior SD, making the initial prior less informative, has little impact on bias detection rates, regardless of choice of $\alpha$.

@@

>>>>

load_path <- here::here("figures","pp15_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","pp10_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","pp6_95CI.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","C95plot.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","simultaneous_sims_plot.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot1.png")
knitr::include_graphics(load_path)

@@

load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")

@@

load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")

@@

load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")

@@

In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ that is selected. For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), we should let $\alpha = 1$ (equal weight between current and historical data). This will maximize the model's ability to detect bias, even when the number of strikes is small. This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes. Higher bias detection rates are obtained for higher values of alpha, acorss compatibility scenarios (Fig X). This effect is most pronounced for scenarios with a low number of strikes and few historical trails (top left paned of fig X). There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X). Overall, though, increasing $\alpha$ will increase bias detection under current and historica trial compatibility.

@@

Finally, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials. This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof. Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias. As @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior.

This sensitivity, however, reduces after including the historical , especially when $\alpha$ is close to 1. Fig X shows the effect of the initial prior’s standard deviation on bias detection rate under a number of compatibility scenarios and alpha values.

@@

As Figure X indicates, if we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data. As alpha increases, there is less of an impact of that lower standard deviation. Moreover, increasing the prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
